---
title: "Case-to-Code: A DNAJC6 Analysis"
author: "Raval Kush"
format: 
  html:
    code-fold: true
    toc: true
    toc-float: true
    theme: cosmo
    embed-resources: true
editor: visual
---

# CASE TO CODE : INFANTILE PARKINSONISM

## 

## \*\*Visualizing the structural defects in DNAJC6 after altered genotype \*\*

```{r}
#| label: Structural and clinical context

# Structural & Clinical Context

# --- 1. Load Libraries ---
library(r3dmol)
library(htmltools)
library(knitr)

# --- 2. Create the Corrected Mutation Table ---
# (This section remains unchanged)
patient_genotypes <- data.frame(
  Sample_Group = c("Control", "CRISPR_P1", "Patient1", "Patient2", "Patient3"),
  Genotype = c(
    "Wild-type",
    "CRISPR-corrected (from Patient 1)",
    "c.766C>T (homozygous)",
    "c.2416C>T (homozygous)",
    "c.801-2 A>G (homozygous)"
  ),
  Mutation_Type = c(
    "Normal",
    "Corrected to WT",
    "Nonsense (Exon 6)",
    "Nonsense (Exon 16)",
    "Splice acceptor (Exon 7)"
  ),
  Predicted_Protein = c(
    "Full-length DNAJC6 (910 aa)",
    "Full-length DNAJC6 (910 aa)",
    "p.R256* - Truncated at residue 256",
    "p.R806* - Truncated at residue 806",
    "Aberrant splicing → likely NMD or exon 7 skip"
  ),
  Functional_Consequence = c(
    "Normal auxilin function",
    "Restored auxilin function",
    "Loss of 72% of protein (clathrin-binding LOST)",
    "Loss of 11% of protein (clathrin-binding LOST)",
    "Likely complete loss of function"
  )
)

knitr::kable(patient_genotypes, 
             caption = "DNAJC6 Genotypes Across Patient Samples",
             align = "l")

# --- 3. Define Key Features & Domains ---
protein_length <- 913   

# Domain coordinates (User-specified)
phosphatase_tensin_domain_start <- 55
phosphatase_tensin_domain_end <- 222
C2_tensin_domain_start <- 228
C2_tensin_domain_end <- 366
j_domain_start <- 849
j_domain_end <- 913

# Mutation locations
mut_p1_residue <- 256 # p.R256*
mut_p2_residue <- 806 # p.R806*

# Exon 7 boundaries (approximate)
exon7_start <- 267   
exon7_end <- 317   


# --- 4. Read PDB Data from Local File ---
pdb_file_path <- "AF-O75061-F1-model_v6.pdb"
pdb_data <- paste(readLines(pdb_file_path), collapse = "\n")


# --- 5. Create Interactive 3D Visualization ---
viewer <- r3dmol(
  width = "100%", 
  height = "600px",
  viewer_spec = m_viewer_spec(
    backgroundColor = "white"
  )
) %>%
  
  # Load the AlphaFold structure
  m_add_model(data = pdb_data, format = "pdb") %>%
  
  # Base style: Color by AlphaFold confidence 
  m_set_style(style = m_style_cartoon(color = "spectrum")) %>%
  
  # --- Highlight Key Functional Domains ---
  
  # Phosphatase Tensin-like Domain (Orange)
  m_set_style(
    sel = list(resi = phosphatase_tensin_domain_start:phosphatase_tensin_domain_end), 
    style = m_style_cartoon(color = "#F39B7F", thickness = 1.2)
  ) %>%
  m_add_label(
    "Phosphatase Tensin-like",
    sel = list(resi = 138),
    style = m_style_label(
      backgroundColor = "#F39B7F", 
      fontColor = "black",
      fontSize = 12
    )
  ) %>%
  
  # C2 Tensin-like Domain (Red)
  m_set_style(
    sel = list(resi = C2_tensin_domain_start:C2_tensin_domain_end), 
    style = m_style_cartoon(color = "#E64B35", thickness = 1.2)
  ) %>%
  m_add_label(
    "C2 Tensin-like Domain",
    sel = list(resi = 297),
    style = m_style_label(
      backgroundColor = "#E64B35", 
      fontColor = "white",
      fontSize = 12
    )
  ) %>%
  
  # J-Domain (Green) - HSC70 interaction
  m_set_style(
    sel = list(resi = j_domain_start:j_domain_end), 
    style = m_style_cartoon(color = "#00A087", thickness = 1.2)
  ) %>%
  m_add_label(
    "J-Domain (HSC70 binding)",
    sel = list(resi = 881),
    style = m_style_label(
      backgroundColor = "#00A087", 
      fontColor = "white",
      fontSize = 12
    )
  ) %>%
  
  # --- Clathrin-Binding Domain Block REMOVED ---
  
  # --- Highlight Patient Mutations ---
  
  # Patient 1: p.R256* (Early truncation - black sphere)
  m_add_style(
    sel = list(resi = mut_p1_residue), 
    style = m_style_sphere(color = "black", radius = 2.0)
  ) %>%
  m_add_label(
    "P1: p.R256* TRUNCATION",
    sel = list(resi = mut_p1_residue), 
    style = m_style_label(
      backgroundColor = "black", 
      fontColor = "white",
      fontSize = 11
    )
  ) %>%
  
  # Show what Patient 1 LOSES (gray out everything after 256)
  m_set_style(
    sel = list(resi = (mut_p1_residue + 1):protein_length), 
    style = m_style_cartoon(color = "lightgray", opacity = 0.3)
  ) %>%
  
  # Patient 3: Exon 7 region (Dark Gray - splice site mutation)
  m_set_style(
    sel = list(resi = exon7_start:exon7_end), 
    style = m_style_cartoon(color = "#4D4D4D", thickness = 1.3)
  ) %>%
  m_add_label(
    "P3: Exon 7 (Splice defect)",
    sel = list(resi = 292), 
    style = m_style_label(
      backgroundColor = "#4D4D4D",
      fontColor = "white",
      fontSize = 11
    )
  ) %>%

  # Patient 2: p.R806* (Late truncation - dark red sphere)
  # This is just BEFORE the J-domain
  m_add_style(
    sel = list(resi = mut_p2_residue), 
    style = m_style_sphere(color = "darkred", radius = 2.0)
  ) %>%
  m_add_label(
    "P2: p.R806* TRUNCATION",
    sel = list(resi = mut_p2_residue), 
    style = m_style_label(
      backgroundColor = "darkred", 
      fontColor = "white",
      fontSize = 11
    )
  ) %>%
  
  # Show what Patient 2 LOSES (the J-domain)
  m_set_style(
    sel = list(resi = (mut_p2_residue + 1):protein_length), 
    style = m_style_cartoon(color = "pink", opacity = 0.4)
  ) %>%

  # Zoom and enable spinning for presentation
  m_zoom_to() %>%
  m_spin(speed = 0.5)

# Display the viewer
viewer

```

## \*\*RStudio Code for Analysis\*\*

### PHASE 1 : Initial Requirements

#### 1.1 : Setting seed and Loading required packages

```{r}
#| label: Setting seed and loading required packages

# 1.1 : Setting seed and Loading required packages

set.seed(123) # Or any integer you like


#install if you havent already
# install.packges()

# For data manipulation (dplyr, tidyr) and plotting (ggplot2)
library(tidyverse) 

#install if you havent already
# if (!require("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")

# For differential expression analysis
library(DESeq2)
library(edgeR)

# For our heatmap
library(pheatmap)

# For GO, KEGG, and GSEA enrichment analysis
library(clusterProfiler) 
library(enrichplot)

#BiocManager::install('EnhancedVolcano')
library(EnhancedVolcano)
library(AnnotationDbi)
library(org.Hs.eg.db) # This is the "dictionary" for human genes

```

### 1.2 : Defining the Analysis Parameters

```{r}
#| label: Defining the Analysis Parameters
# 1.2 : Defining the Analysis Parameters



fdr_cutoff <- 0.05
fc_cutoff <- 2 # Fold Change (not log2)
log2fc_cutoff <- log2(fc_cutoff)
qval_cutoff <- 0.1
```

#### 1.3 : Loading the count data

```{r}
#| label: Loading the count data
#| echo: true

# 1.3 : Loading the count data

# Define the path to your counts file
counts_file_path <- "GSE208353_Raw_gene_counts_matrix.txt" 

# Load the count data
# header = TRUE tells R the first row contains sample names (GSMs)
# row.names = 1 tells R the first column contains the gene names
counts_data <- read.delim(
  file = counts_file_path, 
  header = TRUE, 
  row.names = 1
)

# Always check the first few lines to make sure it loaded correctly
# You should see gene names as row names and GSM IDs as column names
head(counts_data)
# view(counts_data)  #(optional)
```

#### 1.4 : Loading the series matrix data(or metadata)

```{r}
# 1.4 : Loading the series matrix data(or metadata)

# Define the path to your metadata file
metadata_file <- "metadata.csv"

# Load the metadata (comma-separated .csv file)
colData <- read.csv(
  file = metadata_file,
  row.names = 1 # Set the first column (sample_id) as row names
)

# Prepare Metadata for Analysis
# We MUST convert our experimental columns from text to "factors".
# This is how DESeq2 understands our experimental design.
colData$group <- factor(colData$group)
colData$genotype <- factor(colData$genotype)


# Final Check (The "Sanity Check")

# A. Visual Check: Let's look at our objects
print(colData)
str(colData) # You should see 'Factor' for group and genotype

# B. The Critical Match Check:
# This checks if the row names in 'colData' are in the
# EXACT same order as the column names in 'counts_data'.
check_result <- all(rownames(colData) == colnames(counts_data))

# should return true, to verify that the metadata rows matches the counts data columns 

print(check_result)
```

#### 1.5 : Validation 1: The Case

```{r}
# 1.5 : Validation 1: The Case

# CHECKPOINT: Spot-Check DNAJC6 (ENSG00000116675) Counts 

# Define the Ensembl ID for DNAJC6 viz ENSG00000116675
target_gene_id <- "ENSG00000116675"

# Check if the gene exists in our data
target_gene_id %in% rownames(counts_data)

# returns true when the ensembl id you designated is actually there in the rownames column of the counnts_data
  
# Create a data frame for our plot
  dnajc6_plot_df <- data.frame(
    counts = as.numeric(counts_data[target_gene_id, ]),
    group = colData$group
  )
  
# Create the plot using ggplot2
  # We use geom_jitter() to see all the individual dots
  sanity_check_plot <- ggplot(dnajc6_plot_df, aes(x = group, y = counts, color = group)) +
  geom_jitter(size = 3, width = 0.1, show.legend = FALSE) +
  theme_bw(base_size = 14) +
  labs(
    title = "Sanity Check: DNAJC6 (ENSG00000116675)",
    x = "Experimental Group",
    y = "Raw Counts"
  )

# Display the plot
sanity_check_plot


# Check for Nonsense-Mediated Decay (NMD)

# We'll use the dnajc6_plot_df data frame we already made for the sanity check.
# This time, we'll use a boxplot to see the distributions.
dnajc6_counts_boxplot <- ggplot(dnajc6_plot_df, aes(x = group, y = counts, fill = group)) +
  geom_boxplot() +
  geom_jitter(width = 0.1, size = 2) +
  theme_bw(base_size = 14) +
  scale_y_log10() + # Use a log scale to see differences more clearly
  labs(
    title = "DNAJC6 (ENSG00000116675) Expression",
    subtitle = "Comparing Patient 1 (p.R256*), Patient 2 (p.R806*), and Patient 3 (splice-site)",
    x = "Experimental Group",
    y = "Raw Counts (log10 scale)"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(dnajc6_counts_boxplot)

```

#### 1.6 : Validation 2 : Global Quality Control using PCA

```{r}
# 1.6 : Validation 2 : Global Quality Control using PCA

# Create the DESeqDataSet object
# This object bundles our counts, metadata, and experimental design.
# We'll use a simple design formula for this QC step.
dds <- DESeqDataSetFromMatrix(
  countData = counts_data,
  colData = colData,
  design = ~ group 
)

dds <- DESeq(dds)
# Apply variance-stabilizing transformation (vst)
# We MUST do this before making a PCA plot.
# It transforms the raw, noisy count data onto a scale 
# where samples can be fairly compared.
# 'blind = TRUE' is important: it ensures the transformation
# is unbiased by our experimental design.
vsd <- vst(dds, blind = TRUE)

# Generate and print the PCA plot
# We'll use the built-in 'plotPCA' function.
# 'intgroup' tells the plot which metadata columns to use for 
# coloring and shaping the dots. Using both is a great idea.
PCAplot <- plotPCA(vsd, intgroup = c("group", "genotype")) +
    labs(
      title = "PCA of Global Gene Expression",
      subtitle = "Samples clustered by experimental group"
    ) +
    theme_bw(base_size = 14) +
    # This makes the plot axes proportional (looks cleaner)
    coord_fixed() 

PCAplot
```

#### 1.7 : Validation 3 : Global Quality Control using Dispersion Plot

```{r}
# 1.7 : Validation 3 : Global Quality Control using Dispersion Plot

# This checks DESeq2's model fitting. 
# We want to see the black dots (gene estimates) shrinking 
# towards the red line (fitted estimate) with blue outliers flagged.
plotDispEsts(dds) # Use the full, unfiltered dds object before minimal filtering
```

#### 1.8 : Validation 4 : Global Quality Control using sample to sample heatmap

```{r}
# 1.8 : Validation 4 : Global Quality Control using sample to sample heatmap

# This complements the PCA plot. It shows the correlation between samples.
# We use the variance-stabilized data (vsd) we made for the PCA.
library(RColorBrewer) # For heatmap colors

# Calculate distances
sampleDists <- dist(t(assay(vsd))) 
sampleDistMatrix <- as.matrix(sampleDists)

# Define row names based on group
rownames(sampleDistMatrix) <- paste(vsd$group, vsd$genotype, sep="-") 
colnames(sampleDistMatrix) <- NULL # Hide column names for clarity

# Choose colors
colorsQC <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)

# Plot the heatmap
pheatmap(sampleDistMatrix,
         clustering_distance_rows = sampleDists,
         clustering_distance_cols = sampleDists,
         col = colorsQC,
         main = "Sample-to-Sample Distance Heatmap")
```

#### 1.9 : Pre-filtering genes for further analysis

2 methods\
\

```{r}
# 1.9 : Pre-filtering genes for further analysis

# (A) : Pre-filtering Low-Count Genes


# (Trying to replicate what the authors did in their original work and kinda comparing that with the results of a light pre-filter before Deseq-in the next step)

# Our 'dds' object still holds the raw, unfiltered da-ta.
# Now the cpm() function will work.
cpm_mat <- cpm(counts(dds))

# Set the filter criteria
min_cpm <- 1
min_samples <- 3

# keep_genes is a TRUE/FALSE list (op = original paper)
keep_genes_op <- rowSums(cpm_mat >= min_cpm) >= min_samples

# Now, filter the original 'dds' object
dds_filtered_op <- dds[keep_genes_op, ]


# Let's see how many genes we kept
print(paste("Original gene count:", nrow(dds)))
print(paste("Filtered gene count (original work):", nrow(dds_filtered_op)))


# (B) : Minimal Pre-filtering (DESeq2 Best Practice) 


# We will start from our original, full 'dds' object (60,676 genes)
# which we created for the PCA plot.

# Set the minimal filter
# We'll keep genes that have a total of at least 10 reads
# across all 15 samples. This is a very light "junk" filter.
keep <- rowSums(counts(dds)) >= 10

# Filter the dds object
dds_minimal_filtered <- dds[keep, ]


# Let's see how many genes we kept
# This will be *more* genes than our 17,709, and that's good!
print(paste("Original gene count:", nrow(dds)))
print(paste("Filtered gene count (minimal):", nrow(dds_minimal_filtered)))
```

### PHASE 2 : Isogenic Analysis

#### 2.1 : Isogenic Comparision

```{r}
# 2.1 : Isogenic Comparison


# Subset our minimally-filtered dataset
# We take 'dds_minimal_filtered' and keep only the columns
# where the group is "Patient1" or "CRISPR_P1".
dds_isogenic <- dds_minimal_filtered[ , dds_minimal_filtered$group %in% c("Patient1", "CRISPR_P1") ]

# IMPORTANT: Remove unused factor levels
# After subsetting, R still remembers "Patient2", "Patient3", etc.
# This command drops those unused levels, which is crucial for DESeq.
dds_isogenic$group <- droplevels(dds_isogenic$group)

# Set the "Control" Level (Reference)
# We MUST tell DESeq2 our baseline.
# We want to see Patient1 *relative to* CRISPR_P1.
dds_isogenic$group <- relevel(dds_isogenic$group, ref = "CRISPR_P1")

# Run the Analysis
# This single command runs the full DESeq pipeline
# on our 6 selected samples.
dds_isogenic <- DESeq(dds_isogenic)

# Get the Results
# This extracts the results table. DESeq2's automatic
# independent filtering (your friend's point) happens here.
res_isogenic <- results(dds_isogenic)

# CHECK THE RESULTS
# Let's see a summary
# It will tell us how many genes are "up" or "down"
# using the default 10% FDR (p-adj < 0.1).
summary(res_isogenic)

```

#### 2.2 : Visualizations for significant genes : Volcano Plot

```{r}
# 2.2(A) : Volcano Plot 

# Convert our results object to a data frame
# ggplot2 works best with data frames.
# Create Data Frame and Clean IDs
res_isogenic_df <- as.data.frame(res_isogenic) %>%
  rownames_to_column("ensembl_id_version") # Keep original IDs with version

# Clean IDs for mapping (remove version)
res_isogenic_df$ensembl_id_clean <- gsub("\\..*$", "", res_isogenic_df$ensembl_id_version)

# Add a column to identify significant genes
# We'll use the paper's cutoffs: padj < 0.05 and log2FC > 1 as specified before in 1.2
res_isogenic_df <- res_isogenic_df %>%
  mutate(
    significant = case_when(
      padj < fdr_cutoff & abs(log2FoldChange) > log2fc_cutoff ~ "Significant",
      TRUE ~ "Not Significant" # 'TRUE' means "everything else"
    )
  )

# How many genes met the paper's exact cutoff?
sig_genes_cutoff_isogenic <- res_isogenic_df %>%
  filter(significant == "Significant")

print(paste(
  "Total genes meeting paper's cutoff (FDR < 0.05 & FC > 2):", 
  nrow(sig_genes_cutoff_isogenic)
))

# Create the Volcano Plot

isogenic_res_vol <-  ggplot(res_isogenic_df,
    aes(x = log2FoldChange, y = -log10(padj))) +
    geom_point(aes(color = significant), size = 1, alpha = 0.5) +
    scale_color_manual(
      values = c("Not Significant" = "grey", "Significant" = "dodgerblue2" ),
      name = "Significance"
    ) +
    
    # Add vertical lines for our fold-change cutoff
    geom_vline(xintercept = c(-log2fc_cutoff, log2fc_cutoff), linetype = "dashed", color = "black") +
    
    # Add a horizontal line for our p-value cutoff
    geom_hline(yintercept = -log10(fdr_cutoff), linetype = "dashed", color = "black") +
    
    labs(
      title = "Volcano Plot: Patient1 vs. CRISPR_P1",
      x = "log2(Fold Change)",
      y = "-log10(Adjusted P-value)"
    ) +
    theme_minimal(base_size = 14)

isogenic_res_vol




# (B) : Enhanced Volcano Plot 

library(EnhancedVolcano)

# Map Cleaned IDs to Gene Symbols
gene_map_symbols <- mapIds(
    org.Hs.eg.db, 
    keys = res_isogenic_df$ensembl_id_clean, 
    keytype = "ENSEMBL", 
    column = "SYMBOL", 
    multiVals = "first"
    )

# Add Symbol Column to Data Frame 
# Use the cleaned Ensembl ID as the key to look up the symbol
res_isogenic_df$symbol <- gene_map_symbols[res_isogenic_df$ensembl_id_clean]

# Handle NAs: If symbol is NA, use the original Ensembl ID (with version)
res_isogenic_df$symbol <- ifelse(
    is.na(res_isogenic_df$symbol), 
    res_isogenic_df$ensembl_id_version, # Fallback to original ID
    res_isogenic_df$symbol
    )

# Identify Top Genes and DNAJC6 Symbol 

# Find the symbol corresponding to DNAJC6's cleaned ID
dnajc6_symbol <- res_isogenic_df$symbol[res_isogenic_df$ensembl_id_clean == "ENSG00000116675"][1] 

# Get the symbols of the top 10 most significant genes
iso_top_10_symbols <- res_isogenic_df %>%
  filter(!is.na(padj)) %>%
  arrange(padj) %>%
  head(10) %>%
  pull(symbol)

# Combine DNAJC6 and top 10 for labeling
iso_labels_to_show <- unique(c(dnajc6_symbol, iso_top_10_symbols))

# Generate the Plot
isogenic_enh_vol <- EnhancedVolcano(res_isogenic_df,
    lab = res_isogenic_df$symbol, # Use the new symbol column for labels
    x = 'log2FoldChange',
    y = 'padj',
    title = 'Volcano Plot: Patient1 vs. CRISPR_P1',
    subtitle = 'Isogenic Comparison',
    pCutoff = fdr_cutoff,         # Use parameterized cutoff
    FCcutoff = log2fc_cutoff,     # Use parameterized cutoff
    pointSize = 2.0,
    labSize = 4.0,
    selectLab = iso_labels_to_show,   # Label DNAJC6 and top 10 using symbols
    drawConnectors = TRUE,
    widthConnectors = 0.75,
    max.overlaps = Inf,# Ensure all selected labels are shown
    legendPosition = 'right'

    )

isogenic_enh_vol
```

#### 2.3 : Visualizations for top 50 deferentially expressed genes

```{r}
# 2.3 : Heatmap 


# Get Top 50 Genes
iso_top_50_genes <- res_isogenic %>%
  as.data.frame() %>%
  rownames_to_column("ensembl_id") %>%
  filter(!is.na(padj)) %>%
  arrange(padj) %>%
  head(50)

# Get Normalized Counts (for ALL 15 samples)
vsd_counts_all_samples <- assay(vsd)
iso_top_50_counts_all_samples <- vsd_counts_all_samples[iso_top_50_genes$ensembl_id, ]

# Get the 6 sample names we *actually* want to plot
isogenic_sample_names <- colnames(dds_isogenic)

# Filter the counts matrix to only our 6 samples
iso_top_50_counts <- iso_top_50_counts_all_samples[ , isogenic_sample_names]

# Robust Gene Symbol Conversion
gene_map <- mapIds(
  org.Hs.eg.db,
  keys = rownames(iso_top_50_counts),
  keytype = "ENSEMBL",
  column = "SYMBOL"
)
gene_symbols <- ifelse(is.na(gene_map), names(gene_map), gene_map)
rownames(iso_top_50_counts) <- make.names(gene_symbols, unique = TRUE)

# Create Column Annotations
# 'annotation_col' has 6 rows
# 'top_50_counts' now has 6 columns
annotation_col <- data.frame(
  Group = dds_isogenic$group
)
rownames(annotation_col) <- colnames(iso_top_50_counts) # <-- Lengths now match!

# Generate the Improved Heatmap
pheatmap(
  iso_top_50_counts,
  scale = "row",
  cluster_rows = TRUE,
  cluster_cols = TRUE,
  show_rownames = TRUE,
  show_colnames = FALSE,
  main = "Top 50 Genes (Patient1 vs. CRISPR_P1)",
  annotation_col = annotation_col,
  border_color = NA,
  fontsize_row = 8,
)

# The gene symbols are already stored as the row names 
# of the 'top_50_counts' data frame we just plotted.

iso_gene_symbol_list <- rownames(iso_top_50_counts)

# Print the list to the console
print(iso_gene_symbol_list)
```

#### 2.4 : GO Enrichment (Biological Process, Molecular Function and Cellular Compartment)

```{r}
# 2.4 Gene Ontology Enrichment (ORA)

# Get our gene lists
res_isogenic_df <- as.data.frame(res_isogenic) %>%
  rownames_to_column("ensembl_id") # Get Ensembl IDs

iso_all_tested_gene_ids <- rownames(res_isogenic)

iso_sig_gene_ids <- res_isogenic_df %>%
  filter(padj < fdr_cutoff & abs(log2FoldChange) > log2fc_cutoff & !is.na(padj)) %>%
  pull(ensembl_id) # pull() gets the column as a vector


# We'll use gsub() to find a dot (".") and anything after it ("\\..*$")
# and replace it with nothing ("").
iso_all_gene_ids_cleaned <- gsub("\\..*$", "", iso_all_tested_gene_ids)

iso_sig_gene_ids_cleaned <- gsub("\\..*$", "", iso_sig_gene_ids)

# Convert Ensembl IDs to Entrez IDs (using cleaned IDs)
iso_all_entrez_ids <- mapIds(
  org.Hs.eg.db,
  keys = iso_all_gene_ids_cleaned, # Use cleaned IDs
  keytype = "ENSEMBL",
  column = "ENTREZID",
  multiVals = "first"
)

iso_sig_entrez_ids <- mapIds(
  org.Hs.eg.db,
  keys = iso_sig_gene_ids_cleaned, # Use cleaned IDs
  keytype = "ENSEMBL",
  column = "ENTREZID",
  multiVals = "first"
)


# Run the GO Enrichment (for Biological Process)
# to know more use ?enrichGO

iso_go_res_bp <- enrichGO(
  gene = iso_sig_entrez_ids,
  universe = iso_all_entrez_ids,
  OrgDb = org.Hs.eg.db,
  ont = "BP",                 # "BP" = Biological Process
  pAdjustMethod = "BH",
  pvalueCutoff = fdr_cutoff,
  qvalueCutoff = qval_cutoff,
  readable = TRUE             # This converts Entrez IDs back to Symbols
)

# CHECK THE RESULTS
# Let's look at the top 5 hits

#print(head(as.data.frame(go_results_ora), n = 5))

# Plot the Results
# A dot plot is the best way to see this
plot_isogenic_go_res_bp <- print(
  dotplot(iso_go_res_bp, showCategory = 15) + 
    labs(title = "GO Enrichment (ORA): Top 15 Biological Processes") +
    theme_minimal()
)

#print(go_results_ora$Description)

# Run the GO Enrichment (for Molecular funtion)
iso_go_res_mf <- enrichGO(
  gene = iso_sig_entrez_ids,
  universe = iso_all_entrez_ids,
  OrgDb = org.Hs.eg.db,
  ont = "MF",                 # "BP" = Biological Process
  pAdjustMethod = "BH",
  pvalueCutoff = fdr_cutoff,
  qvalueCutoff = qval_cutoff,
  readable = TRUE             # This converts Entrez IDs back to Symbols
)

plot_isogenic_go_res_mf <- print(
  dotplot(iso_go_res_mf, showCategory = 15) + 
    labs(title = "GO Enrichment (ORA): Top 15 Molecular functions") +
    theme_minimal()
)


# Run the GO Enrichment (for Cellular Compartment)
iso_go_res_cc <- enrichGO(
  gene = iso_sig_entrez_ids,
  universe = iso_all_entrez_ids,
  OrgDb = org.Hs.eg.db,
  ont = "CC",                 # "BP" = Biological Process
  pAdjustMethod = "BH",
  pvalueCutoff = fdr_cutoff,
  qvalueCutoff = qval_cutoff,
  readable = TRUE             # This converts Entrez IDs back to Symbols
)

plot_isogenic_go_res_cc <- print(
  dotplot(iso_go_res_cc, showCategory = 15) + 
    labs(title = "GO Enrichment (ORA): Top 15 Cellular Compartments") +
    theme_minimal()
)


```

#### 2.5 : Gene Set Enrichment Analysis

```{r}
# 2.5 : Gene Set Enrichment Analysis (GSEA)


# Use 'select' to get a mapping data frame.
# This is a bit more robust than mapIds for this purpose.
iso_gene_map_df <- AnnotationDbi::select(
  org.Hs.eg.db,
  keys = iso_all_gene_ids_cleaned,
  columns = c("ENTREZID"),
  keytype = "ENSEMBL"
) %>%
  # remove duplicates, just keep first mapping
  distinct(ENSEMBL, .keep_all = TRUE) 

# Join map with our results
# First, add the cleaned ENSEMBL ID to our results
res_isogenic_df$ENSEMBL <- iso_all_gene_ids_cleaned

# Now, join them
iso_res_mapped <- inner_join(res_isogenic_df, iso_gene_map_df, by = "ENSEMBL")


# Handle Duplicate Entrez IDs
# We'll pick the gene with the "strongest" signal (highest absolute stat)
iso_res_ranked <- iso_res_mapped %>%
  filter(!is.na(stat) & !is.na(ENTREZID)) %>%
  # Group by Entrez ID
  group_by(ENTREZID) %>%
  # Find the row with the max absolute 'stat' value
  slice_max(order_by = abs(stat), n = 1) %>%
  ungroup()

# Create the Final Ranked List
# This is the special vector GSEA needs:
# values = the 'stat' column
# names = the 'ENTREZID' column


iso_gene_list_gsea <- iso_res_ranked$stat
names(iso_gene_list_gsea) <- iso_res_ranked$ENTREZID

# CRITICAL: Sort the list in decreasing order
iso_gene_list_gsea_sorted <- sort(iso_gene_list_gsea, decreasing = TRUE)

# Run GSEA for Biological Process
# to know more use ?gseGO
iso_gsea_bp_res <- gseGO(
  geneList = iso_gene_list_gsea_sorted,
  OrgDb = org.Hs.eg.db,
  ont = "BP", # Biological Process
  pvalueCutoff = fdr_cutoff,
  pAdjustMethod = "BH",
  verbose = TRUE # Shows progress
)


# Plot the GSEA Results
# Dot plot of top 20 pathways
iso_gsea_bp_plot <- dotplot(iso_gsea_bp_res, showCategory = 15) +
    labs(title = "GSEA Results: Top 15 Biological Processes") +
    theme_minimal()

iso_gsea_bp_plot



# Run GSEA for Molecular function
iso_gsea_mf_res <- gseGO(
  geneList = iso_gene_list_gsea_sorted,
  OrgDb = org.Hs.eg.db,
  ont = "MF", # Molecular function
  pvalueCutoff = fdr_cutoff,
  pAdjustMethod = "BH",
  verbose = TRUE # Shows progress
)


# Plot the GSEA Results
# Dot plot of top 20 funtions affected
iso_gsea_mf_plot <- dotplot(iso_gsea_mf_res, showCategory = 15) +
    labs(title = "GSEA Results: Top 15 Molecular Functions") +
    theme_minimal()

iso_gsea_mf_plot

# Run GSEA for Cellular Compartment
iso_gsea_cc_res <- gseGO(
  geneList = iso_gene_list_gsea_sorted,
  OrgDb = org.Hs.eg.db,
  ont = "CC", # Molecular function
  pvalueCutoff = fdr_cutoff,
  pAdjustMethod = "BH",
  verbose = TRUE # Shows progress
)


# Plot the GSEA Results
# Dot plot of top 20 compartments
iso_gsea_cc_plot <- dotplot(iso_gsea_cc_res, showCategory = 15) +
    labs(title = "GSEA Results: Top 15 Cellular Compartments") +
    theme_minimal()

iso_gsea_cc_plot

```

#### 2.6 : Counting Keyword Frequencies in GSEA Results

```{r}
# 2.6 : Count Keyword Frequencies in GSEA Results


# Get the list of all descriptions
iso_gsea_descriptions <- as.data.frame(iso_gsea_bp_res)$Description %>%
  c(as.data.frame(iso_gsea_mf_res)$Description) %>%
  c(as.data.frame(iso_gsea_cc_res)$Description)

# Define your keywords of interest (customize this list!)
# We'll use some key themes we've seen.
keywords_to_count <- c(
  # Synaptic Function & Vesicle Recycling
  "synap", "synapse", "synaptic", "synaptogenesis", "vesicle", "clathrin", 
  "endocytosis", "neurotransmitter", "transmission", "recycle", "recycling",

  # Neurodevelopment & Cell Fate
  "develop", "development", "developmental", "neurog", "neurogenesis", "neuron", 
  "neuronal", "differentiation", "fate", "specification", "progenitor", 
  "morphogenesis", "Wnt", "signal", "signaling",

  # Neuronal Structure
  "axon", "axonogenesis", "dendrite", "dendritic", "terminus", "terminal", 
  "cleft", "junction", "growth cone", "projection",

  # Transcription & Regulation
  "transcription", "regulation", "binding", "DNA binding", "protein binding", 
  "factor", "repressor", "activator",

  # Cell-Specific Keywords
  "dopamine", "dopaminergic", "midbrain", "striat", "striatum", "corpus striatum"
)

# Count occurrences for each keyword (case-insensitive)
iso_keyword_counts <- map_int(keywords_to_count, ~sum(str_detect(iso_gsea_descriptions, regex(.x, ignore_case = TRUE))))

# Create a nice summary table
iso_keyword_summary <- tibble(
  Keyword = keywords_to_count,
  Count = iso_keyword_counts
) %>%
  arrange(desc(Count)) # Sort by most frequent

# Print the summary
print(iso_keyword_summary, n = 50) # Print more rows if needed
```

### PHASE 3 : All Patients vs Control

#### 3.1 : Deferential Gene Expression Analysis for all patients vs controls

```{r}
# 3.1 : Analysis 2 - All Patients vs. Controls

# Prepare the Metadata
# We need a new column defining our broad groups: "Patient" vs "Control".
# We'll treat the CRISPR-corrected line as a 'Control' for this comparison,
# as it represents the non-diseased state.
colData <- colData %>%
  mutate(status = case_when(
    group %in% c("Control", "CRISPR_P1") ~ "Control_Group",
    group %in% c("Patient1", "Patient2", "Patient3") ~ "Patient_Group"
  ))

# Convert the new column to a factor
colData$status <- factor(colData$status)

# Create the DESeqDataSet
# We use our 'dds_minimal_filtered' object, which has all 15 samples.
# The design formula is key: ~ genotype + status
# This tells DESeq2: "First, account for differences due to the specific
# 'genotype', and THEN find the differences due to 'status'."
dds_all_patients <- DESeqDataSetFromMatrix(
  countData = counts(dds_minimal_filtered), # Use counts from the filtered object
  colData = colData,                         # Use our updated colData
  design = ~ status
)

# Set the "Control" Level
# We must tell DESeq2 which 'status' is the reference.
dds_all_patients$status <- relevel(dds_all_patients$status, ref = "Control_Group")

# Run the Analysis
# This might take slightly longer as it's using all 15 samples.
dds_all_patients <- DESeq(dds_all_patients)

# Get the Results
# We explicitly ask for the comparison "Patient_Group vs Control_Group".
res_all_patients <- results(dds_all_patients, name = "status_Patient_Group_vs_Control_Group")

# CHECK THE RESULTS
# Let's see the summary (default 10% FDR).
summary(res_all_patients)
```

#### 3.2 : Visualizations for significant genes : Volcano Plot

```{r}
# 3.2 : Volcano Plots

# Setup: Ensure results are in a data frame
# (Assuming res_all_patients is already created)
res_all_patients_df <- as.data.frame(res_all_patients) %>%
  rownames_to_column("ensembl_id_version") # Keep original IDs

res_all_patients_df$ensembl_id_clean <- gsub("\\..*$", "", res_all_patients_df$ensembl_id_version)

# 3.2(A) : Standard ggplot2 Volcano

# Add significance column 
res_all_patients_df <- res_all_patients_df %>%
  mutate(
    significant = case_when(
      padj < fdr_cutoff & abs(log2FoldChange) > log2fc_cutoff ~ "Significant",
      TRUE ~ "Not Significant"
    )
  )

# Count significant genes 
sig_genes_cutoff_all <- res_all_patients_df %>%
  filter(significant == "Significant")

print(paste(
  "Total genes (Analysis 2) meeting cutoff:",
  nrow(sig_genes_cutoff_all)
))

# Create the ggplot Volcano Plot
# (Variables renamed for consistency with Analysis 1)
all_patients_res_vol <- ggplot(res_all_patients_df, aes(x = log2FoldChange, y = -log10(padj))) +
    geom_point(aes(color = significant), size = 1, alpha = 0.5) + 
    scale_color_manual(
      values = c("Not Significant" = "grey", "Significant" = "dodgerblue2"),
      name = "Significance" # Set to match Analysis 1
    ) +
    geom_vline(xintercept = c(-log2fc_cutoff, log2fc_cutoff), linetype = "dashed", color = "black") +
    geom_hline(yintercept = -log10(fdr_cutoff), linetype = "dashed", color = "black") +
    labs(
      title = "Volcano Plot: All Patients vs. Controls",
      x = "log2(Fold Change)",
      y = "-log10(Adjusted P-value)"
    ) +
    theme_minimal(base_size = 14)

print(all_patients_res_vol) # Display the ggplot volcano



# 3.2(B) : EnhancedVolcano Plot with Gene Symbols
library(EnhancedVolcano)

# Map IDs to Symbols for this dataframe
gene_map_symbols_all <- mapIds(
    org.Hs.eg.db, 
    keys = res_all_patients_df$ensembl_id_clean, 
    keytype = "ENSEMBL", 
    column = "SYMBOL", 
    multiVals = "first"
    )

res_all_patients_df$symbol <- gene_map_symbols_all[res_all_patients_df$ensembl_id_clean]
res_all_patients_df$symbol <- ifelse(
    is.na(res_all_patients_df$symbol), 
    res_all_patients_df$ensembl_id_version, 
    res_all_patients_df$symbol
    )

# Identify Top Genes and DNAJC6 Symbol
# (Variables renamed for consistency with Analysis 1)
dnajc6_symbol_all <- res_all_patients_df$symbol[res_all_patients_df$ensembl_id_clean == "ENSG00000116675"][1] 

all_top_10_symbols <- res_all_patients_df %>%
  filter(!is.na(padj)) %>%
  arrange(padj) %>%
  head(10) %>%
  pull(symbol)

all_labels_to_show <- unique(c(dnajc6_symbol_all, all_top_10_symbols))

# Generate the EnhancedVolcano Plot
all_patients_enh_vol <- EnhancedVolcano(res_all_patients_df,
    lab = res_all_patients_df$symbol, 
    x = 'log2FoldChange',
    y = 'padj',
    title = 'Volcano Plot: All Patients vs. Controls',
    subtitle = 'Broader Comparison',
    pCutoff = fdr_cutoff,        
    FCcutoff = log2fc_cutoff,      
    pointSize = 2.0,
    labSize = 4.0,
    selectLab = all_labels_to_show, # Use new consistent variable
    drawConnectors = TRUE,
    widthConnectors = 0.75,
    max.overlaps = Inf,
    legendPosition = 'right'
)


all_patients_enh_vol # Display the plot
```

#### 3.3 : Visualizations for top 50 deferentially expressed genes

```{r}
# 3.3 : Heatmap

# Get Top 50 Genes
all_top_50_genes <- res_all_patients %>%
  as.data.frame() %>%
  rownames_to_column("ensembl_id") %>% # Renamed column to match theme
  filter(!is.na(padj)) %>%
  arrange(padj) %>%
  head(50)

# Get Normalized Counts (for ALL 15 samples)
# We assume 'vsd_counts_all_samples' still exists from the PCA step
# vsd_counts_all_samples <- assay(vsd) 
all_top_50_counts <- vsd_counts_all_samples[all_top_50_genes$ensembl_id, ]

# Robust Gene Symbol Conversion
gene_map_all <- mapIds(
  org.Hs.eg.db,
  keys = rownames(all_top_50_counts), # Using rownames directly
  keytype = "ENSEMBL",
  column = "SYMBOL"
)
gene_symbols_all <- ifelse(is.na(gene_map_all), names(gene_map_all), gene_map_all)
rownames(all_top_50_counts) <- make.names(gene_symbols_all, unique = TRUE)

# Create Column Annotations
# This will have 15 rows for all 15 samples
annotation_col_all <- data.frame(
  Status = colData$status # Using the 'status' column (e.g., "Patient", "Control")
)
rownames(annotation_col_all) <- colnames(all_top_50_counts) 

# Generate the Heatmap
pheatmap(
  all_top_50_counts,
  scale = "row",
  cluster_rows = TRUE,
  cluster_cols = TRUE,
  show_rownames = TRUE,
  show_colnames = FALSE,
  main = "Top 50 Genes (All Patients vs. Controls)",
  annotation_col = annotation_col_all,
  border_color = NA,
  fontsize_row = 8
)

# The gene symbols are stored as the row names 
# of the 'all_top_50_counts' data frame we just plotted.
all_gene_symbol_list <- rownames(all_top_50_counts)

# Print the list to the console
# (Added this section to match your theme)
print(all_gene_symbol_list)
```

#### 3.4 : GO Enrichment (Biological Process, Molecular Function and Cellular Compartment)

```{r}
# 3.4 Gene Ontology Enrichment (ORA)


# Get our gene lists
# We assume 'res_all_patients_df' exists from section 3.2
all_all_tested_gene_ids <- rownames(res_all_patients)

# Get sig genes using the 'significant' column created in 3.2
all_sig_gene_ids <- res_all_patients_df %>%
  filter(significant == "Significant") %>%
  pull(ensembl_id_version) # pull() gets the column as a vector

# Clean the Ensembl IDs (remove version)
all_all_gene_ids_cleaned <- gsub("\\..*$", "", all_all_tested_gene_ids)
all_sig_gene_ids_cleaned <- gsub("\\..*$", "", all_sig_gene_ids)

# Convert Ensembl IDs to Entrez IDs (using cleaned IDs)
all_all_entrez_ids <- mapIds(
  org.Hs.eg.db,
  keys = all_all_gene_ids_cleaned, # Use cleaned IDs
  keytype = "ENSEMBL",
  column = "ENTREZID",
  multiVals = "first"
)

all_sig_entrez_ids <- mapIds(
  org.Hs.eg.db,
  keys = all_sig_gene_ids_cleaned, # Use cleaned IDs
  keytype = "ENSEMBL",
  column = "ENTREZID",
  multiVals = "first"
)

# --- Run GO Enrichment (BP) ---
all_go_res_bp <- enrichGO(
  gene = all_sig_entrez_ids,
  universe = all_all_entrez_ids,
  OrgDb = org.Hs.eg.db,
  ont = "BP",           
  pAdjustMethod = "BH",
  pvalueCutoff = fdr_cutoff,
  qvalueCutoff = qval_cutoff,
  readable = TRUE       
)

# Plot the Results (BP)
plot_all_patients_go_res_bp <- print(
  dotplot(all_go_res_bp, showCategory = 15) + 
    labs(title = "GO Enrichment (ORA - All Patients): Top 15 BP") +
    theme_minimal()
)

# --- Run GO Enrichment (MF) ---
all_go_res_mf <- enrichGO(
  gene = all_sig_entrez_ids,
  universe = all_all_entrez_ids,
  OrgDb = org.Hs.eg.db,
  ont = "MF",           
  pAdjustMethod = "BH",
  pvalueCutoff = fdr_cutoff,
  qvalueCutoff = qval_cutoff,
  readable = TRUE       
)

# Plot the Results (MF)
plot_all_patients_go_res_mf <- print(
  dotplot(all_go_res_mf, showCategory = 15) + 
    labs(title = "GO Enrichment (ORA - All Patients): Top 15 Molecular Functions") +
    theme_minimal()
)

# --- Run GO Enrichment (CC) ---
all_go_res_cc <- enrichGO(
  gene = all_sig_entrez_ids,
  universe = all_all_entrez_ids,
  OrgDb = org.Hs.eg.db,
  ont = "CC",           
  pAdjustMethod = "BH",
  pvalueCutoff = fdr_cutoff,
  qvalueCutoff = qval_cutoff,
  readable = TRUE       
)

# Plot the Results (CC)
plot_all_patients_go_res_cc <- print(
  dotplot(all_go_res_cc, showCategory = 15) + 
    labs(title = "GO Enrichment (ORA - All Patients): Top 15 Cellular Compartments") +
    theme_minimal()
)
```

#### 3.5 : Gene Set Enrichment Analysis

```{r}
# 3.5 : Gene Set Enrichment Analysis (GSEA)

# Use 'select' to get a mapping data frame.
# We assume 'all_all_gene_ids_cleaned' exists from section
all_gene_map_df <- AnnotationDbi::select(
  org.Hs.eg.db,
  keys = all_all_gene_ids_cleaned,
  columns = c("ENTREZID"),
  keytype = "ENSEMBL"
) %>%
  # remove duplicates, just keep first mapping
  distinct(ENSEMBL, .keep_all = TRUE) 

# Join map with our results
# We assume 'res_all_patients_df' 
# Add the cleaned ENSEMBL ID to our results df
res_all_patients_df$ENSEMBL <- all_all_gene_ids_cleaned

# Now, join them
all_res_mapped <- inner_join(res_all_patients_df, all_gene_map_df, by = "ENSEMBL")

# Handle Duplicate Entrez IDs
# We'll pick the gene with the "strongest" signal (highest absolute stat)
all_res_ranked <- all_res_mapped %>%
  filter(!is.na(stat) & !is.na(ENTREZID)) %>%
  # Group by Entrez ID
  group_by(ENTREZID) %>%
  # Find the row with the max absolute 'stat' value
  slice_max(order_by = abs(stat), n = 1) %>%
  ungroup()

# Create the Final Ranked List
all_gene_list_gsea <- all_res_ranked$stat
names(all_gene_list_gsea) <- all_res_ranked$ENTREZID

# CRITICAL: Sort the list in decreasing order
all_gene_list_gsea_sorted <- sort(all_gene_list_gsea, decreasing = TRUE)

# --- Run GSEA for Biological Process ---
all_gsea_bp_res <- gseGO(
  geneList = all_gene_list_gsea_sorted,
  OrgDb = org.Hs.eg.db,
  ont = "BP", # Biological Process
  pvalueCutoff = fdr_cutoff,
  pAdjustMethod = "BH",
  verbose = TRUE # Shows progress
)

# Plot the GSEA Results
all_gsea_bp_plot <- dotplot(all_gsea_bp_res, showCategory = 15) +
    labs(title = "GSEA Results (All Patients): Top 15 Biological Processes") +
    theme_minimal()

all_gsea_bp_plot

# --- Run GSEA for Molecular function ---
all_gsea_mf_res <- gseGO(
  geneList = all_gene_list_gsea_sorted,
  OrgDb = org.Hs.eg.db,
  ont = "MF", # Molecular function
  pvalueCutoff = fdr_cutoff,
  pAdjustMethod = "BH",
  verbose = TRUE # Shows progress
)

# Plot the GSEA Results
all_gsea_mf_plot <- dotplot(all_gsea_mf_res, showCategory = 15) +
    labs(title = "GSEA Results (All Patients): Top 15 Molecular Functions") +
    theme_minimal()

all_gsea_mf_plot

# --- Run GSEA for Cellular Compartment ---
all_gsea_cc_res <- gseGO(
  geneList = all_gene_list_gsea_sorted,
  OrgDb = org.Hs.eg.db,
  ont = "CC", # Cellular Compartment
  pvalueCutoff = fdr_cutoff,
  pAdjustMethod = "BH",
  verbose = TRUE # Shows progress
)

# Plot the GSEA Results
all_gsea_cc_plot <- dotplot(all_gsea_cc_res, showCategory = 15) +
    labs(title = "GSEA Results (All Patients): Top 15 Cellular Compartments") +
    theme_minimal()

all_gsea_cc_plot
```

#### 3.6 : Counting Keyword Frequencies in GSEA Results

```{r}
# 3.6 : Count Keyword Frequencies in GSEA Results (ANALYSIS 2) ###

# Get the list of all descriptions from all three ontologies
# (This now matches the theme)
all_gsea_descriptions <- as.data.frame(all_gsea_bp_res)$Description %>%
  c(as.data.frame(all_gsea_mf_res)$Description) %>%
  c(as.data.frame(all_gsea_cc_res)$Description)

# We assume 'keywords_to_count' is already defined from section 2.6
# keywords_to_count <- c(...) 

# Count occurrences for each keyword (case-insensitive)
all_keyword_counts <- map_int(keywords_to_count, ~sum(str_detect(all_gsea_descriptions, regex(.x, ignore_case = TRUE))))

# Create a nice summary table
all_keyword_summary <- tibble(
  Keyword = keywords_to_count,
  Count = all_keyword_counts
) %>%
  arrange(desc(Count)) # Sort by most frequent

# Print the summary
print(all_keyword_summary, n = 50) # Print more rows if needed
```

### PHASE 4 : Common Analysis

#### 4.1 : Compare Significant Gene Lists

```{r}
# 4.1 : Compare Significant Gene Lists (Aesthetic Venn Diagram)

# Load the libraries
library(ggVennDiagram)
library(ggplot2) 

# 1. Get the significant gene lists (Cleaned IDs)
# We'll use the data frames we created in the volcano plot sections
# (sig_genes_cutoff_isogenic from 2.2 and sig_genes_cutoff_all from 4.2)
iso_sig_ids_cleaned <- sig_genes_cutoff_isogenic$ensembl_id_clean
all_sig_ids_cleaned <- sig_genes_cutoff_all$ensembl_id_clean

# Create the list object
gene_lists <- list(
  Isogenic = iso_sig_ids_cleaned,
 `All Patients` = all_sig_ids_cleaned 
)

# Generate the Aesthetic Venn Diagram
overlap_venn_plot <- ggVennDiagram(
    gene_lists,
    category.names = c("Isogenic Analysis", "All Patients Analysis"),
    label = "both",           # Show count and percent (e.g., 957 (22%))
    label_geom = "text",      # Use plain text for labels (no box)
    label_color = "black",    # Color of the count/percent text
    label_size = 5.5,         # Make count/percent text larger
    label_percent_digit = 1   # Show one decimal place (e.g., 21.5%)
  ) +
  # Use a nice blue/cyan color palette
  scale_fill_gradient(low = "#81D4FA", high = "#0277BD") + 
  
  # Add titles and subtitle
  labs(
    title = "Overlap of Significant Genes",
    subtitle = paste0("FDR < ", fdr_cutoff, " & Fold Change > ", fc_cutoff)
  ) +
  
  # Use a clean, blank theme
  theme_void() + 
  
  # Style and center the titles
  theme(
    plot.title = element_text(hjust = 0.5, size = 18, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 12),
    legend.position = "none" # Hide the gradient legend
  ) +
  
  coord_sf(clip = "off") 

print(overlap_venn_plot) # Display the plot

# Get the list of overlapping genes (The "Core" Signature) 
core_sig_gene_ids <- intersect(iso_sig_ids_cleaned, all_sig_ids_cleaned)

print(paste("Number of core overlapping genes:", length(core_sig_gene_ids)))




```

#### 4.2 : Map Core Signature Genes to Symbols

```{r}
# 4.2 : Map Core Signature Genes to Symbols

# We will use the 'core_sig_gene_ids' vector 
# (from the previous Venn diagram step, section 5.1)

# Map Ensembl IDs to Gene Symbols
# 'core_sig_gene_ids' already contains cleaned IDs (no version).
core_gene_map <- mapIds(
  org.Hs.eg.db,
  keys = core_sig_gene_ids, # Use the consistent variable
  keytype = "ENSEMBL",
  column = "SYMBOL",
  multiVals = "first" # Handle one-to-many mappings
)

# Handle NAs (If a symbol isn't found, keep the Ensembl ID)
core_gene_symbols <- ifelse(
    is.na(core_gene_map), 
    names(core_gene_map), # Use the Ensembl ID as a fallback
    core_gene_map          # Otherwise, use the mapped symbol
    )

# Create a tibble for easier viewing and sorting
# (Using tibble for theme consistency)
core_gene_df <- tibble(
  Ensembl_ID = names(core_gene_map), 
  Gene_Symbol = core_gene_symbols
) %>% 
  arrange(Gene_Symbol) # Sort alphabetically by symbol

# Print the resulting tibble
print(core_gene_df, n = 20) # Print top 20

overlap_df <- data.frame(
  Ensembl_ID = names(core_gene_map), 
  Gene_Symbol = core_gene_symbols
) %>% 
  arrange(Gene_Symbol) # Sort alphabetically by symbol


# You can also print just the list of symbols:
# print(core_gene_df$Gene_Symbol, n = 1000) 

# You might want to save this list to a file for later reference:
#write.csv(core_gene_df, "core_signature_genes.csv", row.names = FALSE)

# You might want to save this list to a file for later reference:
#write.table(overlap_df$Gene_Symbol, "overlapping genes.txt", sep = " ", row.names = FALSE)
```

#### 4.3 : GO Enrichment (Biological Process, Molecular Function and Cellular Compartment)

```{r}
# 4.3 : GO Enrichment (ORA) on Core Signature Genes ###

# Get Gene Lists (Entrez IDs)

# We will use 'core_sig_gene_ids' (Ensembl) from section 4.1
# We will use 'iso_all_entrez_ids' (Entrez) from section 2.4 as our universe

# Convert the Core Signature Ensembl IDs to Entrez IDs
core_sig_entrez_ids <- mapIds(
  org.Hs.eg.db,
  keys = core_sig_gene_ids, # From 5.1
  keytype = "ENSEMBL",
  column = "ENTREZID",
  multiVals = "first"
)

# Run GO Enrichment (BP)
core_go_res_bp <- enrichGO(
  gene = core_sig_entrez_ids,
  universe = iso_all_entrez_ids, # Re-using from 2.4
  OrgDb = org.Hs.eg.db,
  ont = "BP",           
  pAdjustMethod = "BH",
  pvalueCutoff = fdr_cutoff,
  qvalueCutoff = qval_cutoff,
  readable = TRUE       
)

# Plot the Results (BP)
plot_core_go_res_bp <- print(
  dotplot(core_go_res_bp, showCategory = 15) + 
    labs(title = "GO Enrichment (ORA - Core Signature): Top 15 BP") +
    theme_minimal()
)

# Run GO Enrichment (MF)
core_go_res_mf <- enrichGO(
  gene = core_sig_entrez_ids,
  universe = iso_all_entrez_ids, 
  OrgDb = org.Hs.eg.db,
  ont = "MF",           
  pAdjustMethod = "BH",
  pvalueCutoff = fdr_cutoff,
  qvalueCutoff = qval_cutoff,
  readable = TRUE       
)

# Plot the Results (MF)
plot_core_go_res_mf <- print(
  dotplot(core_go_res_mf, showCategory = 15) + 
    labs(title = "GO Enrichment (ORA - Core Signature): Top 15 Molecular Functions") +
    theme_minimal()
)

# Run GO Enrichment (CC)
core_go_res_cc <- enrichGO(
  gene = core_sig_entrez_ids,
  universe = iso_all_entrez_ids, 
  OrgDb = org.Hs.eg.db,
  ont = "CC",           
  pAdjustMethod = "BH",
  pvalueCutoff = fdr_cutoff,
  qvalueCutoff = qval_cutoff,
  readable = TRUE       
)

# Plot the Results (CC)
plot_core_go_res_cc <- print(
  dotplot(core_go_res_cc, showCategory = 15) + 
    labs(title = "GO Enrichment (ORA - Core Signature): Top 15 Cellular Compartments") +
    theme_minimal()
)
```

#### 4.4 : KEGG Analysis

```{r}
# 4.4 : KEGG Enrichment (ORA) on Core Signature Genes

# We will re-use the same Entrez ID lists from section 5.3:
# 'core_sig_entrez_ids' (the core significant genes)
# 'iso_all_entrez_ids' (the background universe from 2.4)

# Run KEGG Enrichment (ORA)
# Note: Requires an internet connection.
core_kegg_res <- enrichKEGG(
  gene = core_sig_entrez_ids,
  universe = iso_all_entrez_ids,
  organism = 'hsa', # 'hsa' is Homo sapiens
  pAdjustMethod = "BH",
  pvalueCutoff = fdr_cutoff,
  qvalueCutoff = qval_cutoff
)

# Convert KEGG IDs back to Gene Symbols (for readability)
if (!is.null(core_kegg_res)) {
  core_kegg_res <- setReadable(core_kegg_res, OrgDb = org.Hs.eg.db, keyType="ENTREZID")
} else {
  print("No significant KEGG pathways found.")
}

# CHECK THE RESULTS
print("--- Top KEGG Pathway Results (ORA - Core Signature) ---")
if (!is.null(core_kegg_res)) {
  print(head(as.data.frame(core_kegg_res), n = 15)) # Show top 15
} 

# Plot the Results (if any were found)
if (!is.null(core_kegg_res) && nrow(as.data.frame(core_kegg_res)) > 0) {
  plot_core_kegg_res <- print(
    dotplot(core_kegg_res, showCategory = 15) + # Changed to 15
      labs(title = "KEGG Enrichment (ORA - Core Signature): Top 15 Pathways") + # Changed to 15
      theme_minimal()
  )
} else {
    print("No significant KEGG pathways to plot.")
}
```

#### 4.5 : Heatmap of Core Signature Genes

```{r}
# 4.5 : Heatmap of Core Signature Genes


# We will use 'core_sig_gene_ids' (Ensembl, clean) from section 4.1
# We will use 'res_isogenic_df' (from 2.2) to get significance ranks
# We will use 'vsd_counts_all_samples' (the full 15-sample count matrix) from 2.1

res_isogenic_df <- as.data.frame(res_isogenic) %>%
  rownames_to_column("ensembl_id_version") # Keep original IDs with version

# Clean IDs for mapping (remove version)
res_isogenic_df$ensembl_id_clean <- gsub("\\..*$", "", res_isogenic_df$ensembl_id_version)

res_isogenic_df$symbol <- gene_map_symbols[res_isogenic_df$ensembl_id_clean]


res_isogenic_df$symbol <- ifelse(
    is.na(res_isogenic_df$symbol), 
    res_isogenic_df$ensembl_id_version, # Fallback to original ID
    res_isogenic_df$symbol
    )

res_isogenic_df <- res_isogenic_df %>%
  mutate(
    significant = case_when(
      padj < fdr_cutoff & abs(log2FoldChange) > log2fc_cutoff ~ "Significant",
      TRUE ~ "Not Significant" # 'TRUE' means "everything else"
    )
  )

# Get the VERSIONED Ensembl IDs for the TOP 50 core genes
top_50_core_versioned_ids <- res_isogenic_df %>%
  # First, keep only the genes from our core signature
  filter(ensembl_id_clean %in% core_sig_gene_ids) %>%
  # Next, sort them by significance (most significant first)
  arrange(padj) %>%
  # Now, take only the top 50
  head(50) %>%
  # Finally, pull their versioned IDs for the count matrix
  pull(ensembl_id_version)

# Get Normalized Counts for these 50 genes across ALL 15 samples
top_50_core_counts <- vsd_counts_all_samples[top_50_core_versioned_ids, ]

# Robust Gene Symbol Conversion
gene_map_top_50_core <- mapIds(
  org.Hs.eg.db,
  keys = rownames(top_50_core_counts),
  keytype = "ENSEMBL",
  column = "SYMBOL"
)
gene_symbols_top_50_core <- ifelse(is.na(gene_map_top_50_core), names(gene_map_top_50_core), gene_map_top_50_core)
rownames(top_50_core_counts) <- make.names(gene_symbols_top_50_core, unique = TRUE)

# Create Column Annotations (for all 15 samples)
# We assume 'colData' is still available
annotation_col_all <- data.frame(
  Status = colData$status # Use the 'status' column (Patient/Control)
)
rownames(annotation_col_all) <- colnames(top_50_core_counts) 

# Generate the Heatmap
pheatmap(
  top_50_core_counts,
  scale = "row",        # Scale genes to see relative changes (Z-score)
  cluster_rows = TRUE,
  cluster_cols = TRUE,
  show_rownames = TRUE, # <-- CHANGED (We can now show the names)
  show_colnames = TRUE, 
  main = "Heatmap of Top 50 Core Signature Genes (All Samples)", # <-- CHANGED
  annotation_col = annotation_col_all,
  border_color = NA,
  fontsize_row = 8      # <-- RE-ENABLED
)
```

### PHASE 5 : Transcription Factor Analysis

#### 5.1 : Upstream Transcription Factor (TF) Enrichment

```{r}
# 5.1 : Upstream Transcription Factor (TF) Enrichment

# Load Libraries
library(decoupleR)
library(gt)        # For aesthetic tables
library(ggrepel)   # For non-overlapping plot labels

# --- 2. Get the Regulon (TF-target list) ---
# We use DoRothEA levels A, B, C for a high-confidence regulon
iso_regulon_dorothea <- get_dorothea(
  organism = "human",
  levels = c('A', 'B', 'C') 
)

# Prepare our Gene List (as a Matrix)
# We re-use 'res_isogenic_df' from section 2.2
iso_tf_input_data <- res_isogenic_df %>%
  # Filter out genes that have no stat or no symbol
  filter(!is.na(stat) & !is.na(symbol) & !grepl("^ENSG", symbol)) %>% 
  dplyr::select(symbol, stat) %>%
  distinct(symbol, .keep_all = TRUE) # Keep one stat per gene symbol

# Convert to a named vector
iso_tf_stat_vector <- iso_tf_input_data$stat
names(iso_tf_stat_vector) <- iso_tf_input_data$symbol

# Coerce to a 1-column matrix (required for run_mlm)
iso_tf_stat_matrix <- as.matrix(iso_tf_stat_vector)
colnames(iso_tf_stat_matrix) <- "isogenic_comparison"

# Run the TF Enrichment (using 'mlm' method)
iso_tf_activities <- run_mlm(
  mat = iso_tf_stat_matrix, 
  network = iso_regulon_dorothea,
  .source = "source",  # TF column in regulon
  .target = "target",  # Target gene column in regulon
  .mor = "mor",        # Mode of regulation (score)
  minsize = 5          # Min 5 target genes per TF
)

# Process and Tidy Results
iso_tf_results <- iso_tf_activities %>%
  # 'source' is the TF, 'condition' is our matrix column name
  dplyr::select(TF = source, Activity_Score = score, P_Value = p_value) %>%
  # Create a BH-adjusted p-value column
  mutate(p_adj = p.adjust(P_Value, method = "BH")) %>%
  arrange(p_adj)

# Display Top TFs (gt Table)
iso_tf_results %>%
  head(20) %>%
  gt() %>%
  tab_header(
    title = "Top 20 Predicted TFs (Isogenic Analysis)",
    subtitle = "Based on multi-linear model (mlm) enrichment"
  ) %>%
  cols_label(
    TF = "Transcription Factor",
    Activity_Score = "Activity Score",
    P_Value = "P-Value",
    p_adj = "Adjusted P-Value"
  ) %>%
  fmt_scientific(
    columns = c(P_Value, p_adj),
    decimals = 2
  ) %>%
  fmt_number(
    columns = Activity_Score,
    decimals = 3
  ) %>%
  data_color(
    columns = Activity_Score,
    palette = c("#3C5488", "white", "#E64B35") # Blue -> White -> Red
  ) %>%
  tab_options(
    table.border.top.color = "black",
    heading.title.font.weight = "bold",
    column_labels.font.weight = "bold"
  )

# --- 7. Visualize Top TFs (Bar Plot) ---
iso_tf_barplot <- iso_tf_results %>%
  head(20) %>%
  mutate(TF = reorder(TF, Activity_Score)) %>%
  ggplot(aes(x = TF, y = Activity_Score, fill = Activity_Score > 0)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  scale_fill_manual(
    values = c("TRUE" = "#E64B35", "FALSE" = "#3C5488"),
    labels = c("Repressed", "Activated"),
    name = "TF Activity"
  ) +
  labs(
    title = "Top 20 Enriched Transcription Factors (Analysis 1)",
    subtitle = "Based on Isogenic (Patient1 vs. CRISPR_P1) DEGs",
    x = "Transcription Factor",
    y = "Activity Score (MLM)"
  ) +
  theme_minimal(base_size = 12) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black")

print(iso_tf_barplot)

# --- 8. Visualize All TFs (Volcano Plot) ---
iso_tf_volcano <- iso_tf_results %>%
  mutate(
    neg_log_p_adj = -log10(p_adj),
    significant = p_adj < fdr_cutoff & abs(Activity_Score) > 0.5 
  ) %>%
  ggplot(aes(x = Activity_Score, y = neg_log_p_adj, color = significant, label = TF)) +
  geom_point(size = 2, alpha = 0.7) +
  scale_color_manual(
    values = c("FALSE" = "gray70", "TRUE" = "#E64B35"),
    name = "Significant"
  ) +
  geom_hline(yintercept = -log10(fdr_cutoff), linetype = "dashed", color = "black") +
  geom_vline(xintercept = c(-0.5, 0.5), linetype = "dashed", color = "black") +
  ggrepel::geom_text_repel(
    data = . %>% filter(p_adj < 0.001 & abs(Activity_Score) > 1), # Label top hits
    size = 3.5,
    max.overlaps = 15
  ) +
  labs(
    title = "TF Activity Landscape (Analysis 1)",
    subtitle = "Significance vs. Effect Size",
    x = "TF Activity Score (MLM)",
    y = "-log10(Adjusted P-value)"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")

print(iso_tf_volcano)
```

#### 5.2 Analysis of Key TFs

```{r}
# 5.2 : Analysis of Key TFs

# Check Key Neurodevelopmental TFs
neuro_tfs_to_check <- c(
  "FOXA2", "LMX1A", "LMX1B", "NR4A2", "PITX3", "EN1", "EN2", 
  "SOX2", "PAX6", "NEUROD1", "TBR1", "REST", "FOXO1", "CREB1", "MEF2C",
  "EGR1", "CTCF", "ATF4" # Adding TFs from your report
)

iso_neuro_tf_results <- iso_tf_results %>%
  filter(TF %in% neuro_tfs_to_check) %>%
  arrange(p_adj) %>%
  mutate(
    Direction = ifelse(Activity_Score > 0, "Activated", "Repressed"),
    Significant = ifelse(p_adj < fdr_cutoff, "Yes", "No")
  )

# Display with gt
iso_neuro_tf_results %>%
  gt() %>%
  tab_header(title = "Activity of Key Neurodevelopmental TFs") %>%
  fmt_scientific(columns = P_Value, decimals = 2) %>%
  fmt_scientific(columns = p_adj, decimals = 2) %>%
  fmt_number(columns = Activity_Score, decimals = 3) %>%
  data_color(
    columns = Activity_Score,
    palette = c("#3C5488", "white", "#E64B35")
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = Significant,
      rows = Significant == "Yes"
    )
  )

# Check if Significantly Active TFs are also DEGs
iso_tf_significant_names <- iso_tf_results %>%
  filter(p_adj < fdr_cutoff) %>%
  pull(TF)

iso_tf_expression <- res_isogenic_df %>%
  filter(symbol %in% iso_tf_significant_names) %>%
  dplyr::select(TF = symbol, log2FoldChange, padj) %>%
  mutate(
    Status = ifelse(padj < fdr_cutoff, "Significant DE", "Not DE")
  ) %>%
  arrange(padj)

# Display with gt
iso_tf_expression %>%
  gt() %>%
  tab_header(title = "Expression Status of Significantly Active TFs") %>%
  fmt_number(columns = log2FoldChange, decimals = 2) %>%
  fmt_scientific(columns = padj, decimals = 2) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = Status,
      rows = Status == "Significant DE"
    )
  ) %>%
  tab_footnote(
    footnote = "Note: Activity can change (e.g., via phosphorylation) even if the TF's own gene expression does not.",
    locations = cells_title()
  )

# 3. Export Results to File
write.csv(
  iso_tf_results %>% filter(p_adj < fdr_cutoff),
  "TF_enrichment_significant_results.csv",
  row.names = FALSE
)
```

### PHASE 6 : STRING DB

<https://version-12-0.string-db.org/cgi/network?networkId=bAVOdXXjJ2xe>

#### 6.1 : STRING DB Network

![](string%20db%20overlapping%20genes/stringdb_overlapping_genes_vector_graphic.svg)

#### 6.2 : Top enrichment processes

![](string%20db%20overlapping%20genes/stringdb_overlapgenes_enrichment_Process_sim0.8_graph.svg)

### PHASE 7 : Identifying potential lncRNA and Non-coding gene involvement

#### 7.1 : Novel lncRNA & Non-Coding Gene Discovery

```{r}
# 7.1 : Novel Non-Coding Gene Discovery

# Load Libraries
library(biomaRt)
library(DT) # For interactive tables

# Get All Significant Gene IDs (from Isogenic Analysis)
# We re-use 'res_isogenic_df' from section 2.2
iso_sig_genes_df <- res_isogenic_df %>%
  filter(significant == "Significant") # Use the column we already made

iso_sig_ensembl_cleaned <- iso_sig_genes_df$ensembl_id_clean

# Find "Unmappable" IDs (the non-coding candidates)
# Find genes that do NOT map to an Entrez ID (which org.Hs.eg.db uses)
entrez_map <- mapIds(org.Hs.eg.db,
                     keys = iso_sig_ensembl_cleaned,
                     keytype = "ENSEMBL",
                     column = "ENTREZID",
                     multiVals = "first")

iso_unmappable_ids <- names(entrez_map[is.na(entrez_map)])

print(paste("Found", length(iso_unmappable_ids), "significant genes not in Entrez/org.Hs.eg.db (non-coding candidates)"))

# Annotate Unknown IDs with biomaRt
if (length(iso_unmappable_ids) > 0) {
  ensembl_mart <- useMart(biomart = "ensembl", dataset = "hsapiens_gene_ensembl")
  
  iso_gene_annotations <- getBM(
    attributes = c('ensembl_gene_id', 'external_gene_name', 'gene_biotype', 'description'),
    filters = 'ensembl_gene_id',
    values = iso_unmappable_ids, 
    mart = ensembl_mart
  )

  # Join with expression data
  iso_novel_hits_df <- iso_gene_annotations %>%
    # Join annotations with our DESeq2 results
    left_join(
      res_isogenic_df %>% dplyr::select(ensembl_id_clean, log2FoldChange, padj, baseMean), 
      by = c("ensembl_gene_id" = "ensembl_id_clean")
    ) %>%
    distinct(ensembl_gene_id, .keep_all = TRUE) %>%
    arrange(padj) # Sort by most significant

  # --- 6. Display All Novel Hits (Interactive) ---
  print("--- Discovered Significant Non-Coding & Novel Genes ---")
  print(
    datatable(iso_novel_hits_df, 
              caption = "Top Dysregulated Non-Coding & Unmapped Genes",
              options = list(pageLength = 10, scrollX = TRUE),
              rownames = FALSE,
              filter = 'top'
              )
  )
  
  # Save the list
  write.csv(iso_novel_hits_df, "novel_significant_genes.csv", row.names = FALSE)

  # Filter for and Display lncRNAs
  iso_lncrna_hits_df <- iso_novel_hits_df %>%
    filter(gene_biotype == "lncRNA")
    
  print("--- Of these, the following are lncRNAs: ---")
  print(
    datatable(iso_lncrna_hits_df, 
              caption = "Top Dysregulated lncRNAs",
              options = list(pageLength = 10, scrollX = TRUE),
              rownames = FALSE,
              filter = 'top'
              )
  )
  
  write.csv(iso_lncrna_hits_df, "novel_significant_lncRNAs.csv", row.names = FALSE)
  
} else {
  print("No unmapped significant genes were found.")
}
```

#### 7.2 : Co-expression Heatmap

```{r}
# 7.2 : Co-expression Heatmap (lncRNA vs. Key Genes) 

# Load Libraries
library(pheatmap)

# Define Target Gene Lists
# We re-use 'iso_lncrna_hits_df' from 7.1
iso_top_lncrnas <- iso_lncrna_hits_df %>%
  head(5) %>% # Take top 5 most significant
  pull(external_gene_name) # Get their gene symbols

# Key PD & Neurodevelopmental genes to check against
iso_key_dev_genes <- c(
  "DNAJC6", "SNCA", "CLTC",  # Key PD/CME genes
  "DLX1", "DLX2", "DLX5", "DLX6", # DLX family (partner for DLX6-AS1)
  "FOXA2", "NEUROD1", "LMX1A", "PAX6", # Top neurodev TFs from ORA
  "EGR1", "ATF4", # Top TFs from decoupleR
  "DNAJC6", # Your causal gene
  "SNCA",   # Alpha-synuclein
  "CLTC",   # Clathrin Heavy Chain
  "SYN1"    # Synapsin I
)

iso_all_heatmap_symbols <- unique(c(iso_top_lncrnas, iso_key_dev_genes))

# Get Normalized Counts for Target Genes
# We re-use 'vsd' (from 2.1) and 'res_isogenic_df' (from 2.2)
if (!exists("vsd")) {
  stop("Error: 'vsd' object not found. Please run the PCA chunk (2.1) first.")
}

# Find the Ensembl IDs (with version) that match our symbols
iso_target_ensembl_ids <- res_isogenic_df %>%
  filter(symbol %in% iso_all_heatmap_symbols) %>%
  distinct(symbol, .keep_all = TRUE) %>% # Handle duplicate symbols
  pull(ensembl_id_version)
  
# Get normalized counts for these genes across all 15 samples
# and transpose it (so genes are columns, samples are rows)
iso_cor_matrix_data <- t(assay(vsd)[iso_target_ensembl_ids, ])

# Set Column Names to Gene Symbols
# We need to map the versioned IDs back to symbols
symbol_map_df <- res_isogenic_df %>%
  filter(ensembl_id_version %in% iso_target_ensembl_ids) %>%
  dplyr::select(ensembl_id_version, symbol)

# Match the column names (versioned IDs) to their symbols
colnames(iso_cor_matrix_data) <- symbol_map_df$symbol[
  match(colnames(iso_cor_matrix_data), symbol_map_df$ensembl_id_version)
]

# Calculate Correlation Matrix
iso_cor_matrix <- cor(iso_cor_matrix_data, method = "pearson", use = "pairwise.complete.obs")

# Plot Correlation Heatmap
print(
  pheatmap(
    iso_cor_matrix,
    main = "Co-expression: Top lncRNAs vs. Key PD & Development Genes",
    fontsize_row = 10,
    fontsize_col = 10,
    color = colorRampPalette(c("#3C5488", "white", "#E64B35"))(50), # Blue-White-Red
    cluster_rows = TRUE,
    cluster_cols = TRUE,
    display_numbers = TRUE, # Show correlation values
    number_format = "%.2f", # Format to 2 decimal places
    fontsize_number = 8,
    border_color = "grey60"
  )
)
```

```{r}
### Session Information ###
sessionInfo()
```
